---
title: "DS311 - R Lab Assignment"
author: "Barbara Wallen"
date: "8/22/2022"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
ncol(mtcars)
nrow(mtcars)

# Answer:
print("There are total of __32___ variables and __352___ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are __6___ discrete variables and __5___ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!

m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sqrt(v)


 print(paste("The average of Mile Per Gallon from this data set is ", m  , " with variance ", v , " and standard deviation", s  , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!

library(dplyr)
library(tidyr)

mtcars %>%
  group_by(cyl, gear) %>%
  tally() %>%
  spread(cyl, n)

avgclass = mtcars %>% group_by(mtcars$cyl) %>% summarise(mean_mpg = mean(mtcars$mpg))
print(avgclass)
```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!


print("The most common car type in this data set is car with ____ cylinders and ____ gears. There are total of _____ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
hist <- PlantGrowth$weight
hist(hist,
     main = 'Plant Growth',
     xlab = 'weight',
     ylab = 'frequency' )

```

Result:

=> Report a paragraph to summarize your findings from the plot!

There is 8 plants with the weight between 5 to 5.5 and 2 plants with weight between 3.5 to 4.0


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
miles <- mtcars$mpg 
hist(miles,
     breaks = 10,
     main = 'MPG',
     xlab = 'miles per gallon',
     ylab = 'number of cars' )

print("Most of the cars in this data set are in the class of __7____ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
library(ggplot2)
plot(y= USArrests$Murder, x= USArrests$Assault, main = "Murder and Assaults Rate")

```

Result:

=> Report a paragraph to summarize your findings from the plot!

The ratio between murders is lower than assaults, meaning that assaults has a high rate

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!

str(housingData)
min(housingData$Market.Value.per.SqFt)
max(housingData$Market.Value.per.SqFt)
mean(housingData$Market.Value.per.SqFt)

min(housingData$Year.Built)
max(housingData$Year.Built)
mean(housingData$Year.Built)

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
avg <- aggregate(housingData$Market.Value.per.SqFt, list(housingData$Boro), FUN=mean) 
avg
barplot(matrix(avg$x), beside=T, main='Average price', names.arg = avg$Group.1, ylab="Average prices", xlab="Boro")



```

c. Write a summary about your findings from this exercise.
The houses from Manhattan is the most expensive (180k) from this data set, and the cheapest house ($41k) is found on State Island neighborhood.  

The oldest house is from 1825.

=> Enter your answer here!


